\thetitle

### 2020-11-18 sun_phases/spin_qudit_tomo.tex
overhaul of the reconstruction error bound calculation

Consider a $d$-state spin qudit with total spin $s\equiv\frac{d-1}{2}$.
The defining property of a spin qudit is the fact that it describes an angular momentum degree of freedom, which has specific implications for how a spin qudit should transform under the group SO(3) of rotations in 3D space.
Due to the central importance of these transformation rules for a spin qudit, we first identify a basis of operators that transforms nicely under 3D rotations.
One such basis is provided by the {\it transition operators} (also known as polarization operators \cite{kryszewski2006positivity, bertlmann2008bloch}), defined by
\begin{align}
  T_{\ell m} \equiv \sqrt{\f{2\ell+1}{2s+1}} \sum_{\mu,\nu=-s}^s
  \bk{s\mu;\ell m|s\nu} \op{\nu}{\mu},
  \label{eq:trans_op}
\end{align}
where $\bk{s\mu;\ell m|s\nu}$ is a Clebsh-Gordan coefficient that enforces $\ell\in\set{0,1,\cdots,d-1}$ and $m\in\set{-\ell,-\ell+1,\cdots,\ell}$, such that there are $d^2$ operators in total.
For brevity, we will generally treat the value of $d$ (or $s$) as constant but arbitrary throughout this work, and suppress any explicit dependence of quantities or operators such as $T_{\ell m}$ on $d$.
The transition operators are orthonormal with respect to the trace inner product, and transform nicely under conjugation:
\begin{align}
  \obk{T_{\ell m }|T_{\ell'm'}}
  = \delta_{\ell\ell'} \delta_{mm'},
  &&
  T_{\ell m}^\dag = \p{-1}^m T_{\ell,-m},
\end{align}
where for any $d\times d$ matrix $X=\sum_{\mu,\nu} X_{\mu\nu} \op{\mu}{\nu}$ we define the $d^2$-component vector $\oket{X} \equiv \sum_{\mu,\nu} X_{\mu\nu} \ket{\mu\nu}$; $\obra{X}$ is the conjugate transpose of $\oket{X}$, such that $\obk{X|Y}=\tr\p{X^\dag Y}$; and $\delta_{kk'}\equiv 1$ if $k=k'$ and $0$ otherwise.
These properties of the transition operators allow us to expand any density operator $\rho$ in the transition operator basis as
\begin{align}
  \rho = \sum_{\ell=0}^{d-1} \sum_{m=-\ell}^\ell
  \rho_{\ell m} T_{\ell m}^\dag,
  &&
  \rho_{\ell m} \equiv \bk{T_{\ell m}}_\rho,
  \label{eq:trans_state}
\end{align}
where $\bk{X}_\rho\equiv\obk{\rho|X}=\tr\p{\rho X}$, and $\rho^\dag=\rho$ implies that $\rho_{\ell m}^*=\p{-1}^m\rho_{\ell,-m}$.
The transition operators can be physically interpreted as the transition $\ket\psi\to T_{\ell m}\ket\psi$ induced on the state $\ket\psi$ of a spin-$s$ system by the absorption of a spin-$\ell$ particle with spin projection $m$ onto a fixed quantization axis.
In analogy with the spherical harmonics $Y_{\ell m}$, we will refer to the index $\ell$ as the {\it degree} and $m$ as the {\it order} of $T_{\ell m}$.

@@

Reconstructing an arbitrary density operator $\rho$ essentially requires us to find a set of axes $V=\set{\v v}$ and associated coefficients $C_{\ell mk}\p{\v v}$ that would allow us to recover any coefficient $\rho_{\ell m}$ in expansion in Eq.~\eqref{eq:trans_state} through
\begin{align}
  \rho_{\ell m} = \bk{T_{\ell m}}_\rho
  = \sum_{\v v,k} C_{\ell mk}\p{\v v} \bk{T_{\v vk,0}}_\rho.
  \label{eq:state_recon}
\end{align}
Expanding the rotated transition operators $T_{\v vk,0}$ into a sum of un-rotated transition operators, according to Eq.~\eqref{eq:trans_rot}, we find that the recovery condition in Eq.~\eqref{eq:state_recon} is satisfied when
\begin{align}
  T_{\ell m}
  = \sum_{\v v,k,n} C_{\ell mk}\p{\v v} D^k_{n,0}\p{\v v} T_{kn}.
\end{align}
Orthogonality of the transition operators motivates the ansatz $C_{\ell mk}\p{\v v}=\delta_{k\ell}C_{\ell m}\p{\v v}$, which in turn implies that
\begin{align}
  \sum_{\v v} C_{\ell m}\p{\v v} D^\ell_{n,0}\p{\v v} = \delta_{mn}
  \label{eq:tomo_recovery}
\end{align}
for all $\ell$.

In fact, the problem of finding suitable axes $V$ and coefficients $C_{\ell m}\p{\v v}$ to satisfy Eq.~\eqref{eq:tomo_recovery}, solving which would enable the recovery of arbitrary spin qudit states, can be mapped onto the well-studied problem of signal recovery on the sphere \cite{mcewen2011novel, rauhut2011sparse, alem2012sparse, khalid2014optimaldimensionality}.
The signal recovery problem can be stated as follows: given a square-integrable function $f$ on the sphere, with the harmonic expansion
\begin{align}
  f\p{\v v} = \sum_{\ell,m} f_{\ell m} Y_{\ell m}\p{\v v},
\end{align}
where $f_{\ell m}$ are expansion coefficients and $Y_{\ell m}$ is the spherical harmonic of degree $\ell$ and order $m$, find a set of points $V=\set{\v v}$ and associated coefficients $\tilde C_{\ell m}\p{\v v}$ with which we can reconstruct $f$, or equivalently its coefficients $f_{\ell m}$, from knowledge of the function's value $f\p{\v v}$ at all points $\v v\in V$; that is
\begin{align}
  f_{\ell m} = \sum_{\v v} \tilde C_{\ell m}\p{\v v} f\p{\v v}
  = \sum_{\v v,k,n} \tilde C_{\ell m}\p{\v v} Y_{kn}\p{\v v} f_{kn}.
\end{align}
Reconstruction of functions with arbitrary coefficients $f_{\ell m}$ implies that
\begin{align}
  \sum_{\v v} \tilde C_{\ell m}\p{\v v} Y_{kn}\p{\v v}
  = \delta_{k\ell} \delta_{mn},
  \label{eq:full_recovery}
\end{align}
which is a stronger version of the condition that we found for the spin qudit tomography problem in Eq.~\eqref{eq:tomo_recovery}.
We will refer to Eq.~\eqref{eq:full_recovery} as the {\it full recovery problem}, and Eq.~\eqref{eq:tomo_recovery} as the the {\it reduced recovery problem}.
Due to the fact that $D^k_{n,0}\p{\v v} = \sqrt{\frac{4\pi}{2\ell+1}}\, Y_{kn}\p{\v v}^*$, any solution to the full recovery problem automatically solves the reduced recovery problem by setting $C_{\ell m}\p{\v v} = \sqrt{\frac{2\ell+1}{4\pi}}\, \tilde C_{\ell m}\p{\v v}^*$.
In principle, this mapping allows us to import a host of existing signal recovery algorithms \cite{mcewen2011novel, rauhut2011sparse, alem2012sparse, khalid2014optimaldimensionality} for the task of spin qudit tomography.
In practice, spin qudits typically have only a modest dimension $d$, which allows for simpler and optimized tomography protocols that are practical despite worse scaling with $d$ (see Section \ref{sec:protocol}).

@@

Since
\begin{enumerate*}
\item there are only $2\ell+1\le2d-1$ transition operators of degree $\ell\le\ell_{\t{max}}$, and
\item measuring spin projection along any single axis provides measurement data for transition operators of {\it all} degrees $\ell\le\ell_{\t{max}}$,
\end{enumerate*}
one might expect that $2d-1$ suitably chosen measurement axes would always suffice to perform full tomography of a $d$-level spin qudit.
In the language of classical signal recovery, we would expect there to exist a fixed set of $2\ell_{\t{max}}+1$ points on the sphere that enable the recovery of {\it any} function with fixed degree $\ell\le\ell_{\t{max}}$ (given prior knowledge of $\ell$) \footnote{We could further conjecture that if {\it any} set of points $V\subset\Omega_2$ enables the recovery of degree-$\ell_{\t{max}}$ functions on the sphere, then $V$ would allow recovery degree-$\ell$ functions with $\ell\le\ell_{\t{max}}$.  However, this condition is stronger than that necessary to guarantee full recovery of a $d$-level spin qudit state with $2d-1$ measurement axes.}.
Unfortunately, we have no general proof of this conjecture, which requires a deeper analysis of fixed-degree signal recovery on the sphere.
Nonetheless, in the following section we numerically verify that $2d-1$ measurement axes suffice for all $d\le200$ [\red{todo: possibly refer to Section \ref{sec:axes} instead}].

@@

Computing the classical error scale $\S_V^2$ to certify the bound in Eq.~\eqref{eq:bound_eps} requires constructing the measurement matrix $M_V$ and finding its singular values.
The complexity of this task can be greatly reduced by the fact that the degree $\ell$ of a transition operator $T_{\ell m}$ is preserved under rotations, which implies that the unitary
\begin{align}
  U \equiv \sum_{\ell=0}^{d-1} \sum_{m=-\ell}^\ell
  \oket{T_{\ell m}} \bra{\ell m},
  \label{eq:trans_basis}
\end{align}
with vectors $\oket{T_{\ell m}}$ in a column indexed by integers $\p{\ell,m}$, block-diagonalizes the measurement matrix into $d$ blocks indexed by the degree $\ell$:
\begin{align}
  U^\dag M_V = \sum_{\ell=0}^{d-1} \op{\ell} \otimes M_{V\ell},
\end{align}
where the $\p{2\ell+1}\times\abs{V}$ blocks are
\begin{align}
  M_{V\ell} \equiv \sum_{m,\v v} \ket{m} \obk{T_{\ell m}|T_{\v v\ell,0}} \bra{\v v}
  = \sum_{m,\v v} D^\ell_{m,0}\p{\v v} \op{m}{\v v}.
\end{align}
As the singular values of $M_V$ are invariant under unitary transformations, it follows that
\begin{align}
  \S_V^2 = \sum_\ell \S_{V\ell}^2,
  &&
  \S_{V\ell}^2 \equiv \norm{M_{V\ell}^{-1}}^2,
\end{align}
where $M_{V\ell}^{-1}$ is the right inverse of $M_{V\ell}$.
Constructing the block $M_{V\ell}$ and computing its singular values takes at most $O(\abs{V}d^2)$ time (see Appendix \ref{sec:compute} for details on efficiently computing $\S_{V\ell}^2$).
If we assume that $\abs{V}\sim d$, then computing $\S_V^2$ to certify the bound in Eq.~\eqref{eq:bound_eps} takes $O(d^4)$ serial or $O(d^3)$ parallel runtime (see Figure \ref{fig:times}).

\begin{figure}
  \centering
  \includegraphics{error_scale/error_times.pdf}
  \caption{Serial runtime to compute $\S_V$ with $\abs{V}=2d-1$ randomly chosen measurement axes, averaged over 100 calculations of $\S_V$ or 5 minutes of runtime (for each $d$), whichever comes first.
    Dashed line shows a fit to runtime $t=c d^\alpha$ for all $d\ge100$, finding $\alpha\approx 3.7$.
    These results do not include fixed $O(d^3)$ runtimes for pre-computing quantities that can be recycled for every new choice of measurement axes $V$.}
  \label{fig:times}
\end{figure}

The assumption that observables can be estimated up to uncorrelated noise with maximal variance $\epsilon^2$, summarized by Eq.~\eqref{eq:error_assumption}, is reasonable when measurement error is dominated by classical sources of experimental noise.
However, this assumption breaks down when measurement error is limited by fundamental quantum (shot) noise.
We relax the assumption of Eq.~\eqref{eq:error_assumption} in Appendix \ref{sec:error}, where we instead assume that $\tilde\rho_V$ is built from $n$ independent measurements of spin projection along every axis $\v v\in V$, with shot noise the dominant source of error.
In this case, the mean squared distance between the approximation $\tilde\rho_V$ and the true state $\rho$ can be compactly expressed using trace inner products of the {\it covariance matrices}
\begin{align}
  \C_\ell\sp{\rho} \equiv \sum_{m,n=-\ell}^\ell
  \cov_\rho\p{T_{\ell m}^\dag, T_{\ell n}} \op{m}{n}
\end{align}
where $\cov_\rho\p{X,Y} \equiv \bk{XY}_\rho - \bk{X}_\rho \bk{Y}_\rho$, and the {\it noise matrices}
\begin{align}
  \N_{V\ell} \equiv M_{V\ell} \diag\sp{M_{V\ell}^{-1}
    \p{M_{V\ell}^{-1}}^\dag} M_{V\ell}^\dag,
\end{align}
where $\diag\sp{X}$ sets all off-diagonal parts of $X$ to zero:
\begin{align}
  \Bbk{\norm{\tilde\rho_V-\rho}^2}
  = \f1n \sum_{\ell>0} \obk{\N_{V\ell}|\C_\ell\sp{\rho}}.
  \label{eq:error}
\end{align}
Maximizing this error over all $\rho$ (for a fixed set of axes $V$) yields a reconstruction error bound of the form
\begin{align}
  \Bbk{\norm{\tilde\rho_V-\rho}^2} \le \f{\epsilon_V^2}{n},
  \label{eq:bound}
\end{align}
where the complexity of computing the {\it quantum error scale} $\epsilon_V^2$ (defined below) is the same as that of computing the classical error scale $\S_V^2$ (see Figure \ref{fig:times}).
For comparison with the bound in Eq.~\eqref{eq:bound_eps}, we note that $\epsilon_V^2<\S_V^2/2$, so the previous bound still holds with the replacement $\epsilon^2\to1/2n$.
We also note that the (squared) reconstruction error in Eq.~\eqref{eq:error} and its bound in Eq.~\eqref{eq:bound} obey the ``standard quantum limit'' of $\sim1/\sqrt{n}$ scaling in the number of measurements.
In principle, this scaling can be improved to $\sim1/n$ by preparing and measuring entangled copies of many qudits \cite{giovannetti2006quantum}.

@@

The remainder of this section is devoted to defining $\epsilon_V^2$, and can be skipped without consequence for the discussions in the rest of this paper.
To write out the quantum error scale $\epsilon_V^2$ explicitly, we define a matrix of structure constants
\begin{align}
  g_{L\ell} \equiv \sum_{m,n} g^L_{\ell mn} \op{m}{n},
  &&
  g^L_{\ell mn} \equiv \obk{T_{L,m+n}|T_{\ell m}T_{\ell n}},
\end{align}
and the inversion operator $\I\equiv\sum_m\p{-1}^m\op{m}{m}$, which define components of the $d^2$-component vector
\begin{align}
  \ket{\chi_V}
  &\equiv \sum_{L=0}^{d-1} \sum_{M=-L}^L \chi^V_{LM} \ket{LM}, \\
  \chi^V_{LM} &\equiv \sum_\ell \obk{\N_{V\ell}|\D_M|\I g_{L\ell}}, \\
  \D_M &\equiv \sum_{m,n} \delta_{M,m+n} \op{mn}.
\end{align}
We then define the state vector
\begin{align}
  \oket{\rho_V^\star} = \f1d \oket{\1_d} + U \sum_{\ell>0} \sp{\N_{V\ell} + \I\N_{V\ell}^*\I}^{-1} \ket{\chi_{V\ell}},
\end{align}
where $\1_d$ is the $d\times d$ identity matrix, $\ket{\chi_{V\ell}} \equiv \sum_{m=-\ell}^\ell \ket{\ell m} \bk{\ell m|\chi_V}$ is the projection of $\ket{\chi_V}$ onto the subspace of fixed degree $\ell$, the unitary $U$ defined in Eq.~\eqref{eq:trans_basis} takes $\ket{\ell m}\to \oket{T_{\ell m}}$, and $\N_{V\ell}^*$ is the complex conjugate of $\N_{V\ell}$.
The quantum error scale $\epsilon_V^2$ is then
\begin{align}
  \epsilon_V^2 \equiv \sum_{\ell>0} \obk{\N_{V\ell}|\C_\ell\sp{\rho_V^\star}}.
  \label{eq:scale}
\end{align}
We derive Eq.~\eqref{eq:scale} in Appendix \ref{sec:error}, and discuss its efficient calculation in Appendix \ref{sec:compute}.
Note that the ``state'' $\rho_V^\star$, which maximizes the reconstruction error in Eq.~\eqref{eq:error}, has trace $\tr\rho_V^\star=1$, but may generally have negative eigenvalues.
In principle, there is therefore room for improvement on the bound in Eq.~\eqref{eq:bound} by maximizing reconstruction error over the (restricted) domain of states $\rho$ with a non-negative spectrum.

@@

The ability to certify a statistical error bound on the empirical estimate $\tilde\rho_V$ of an unknown quantum state $\rho$ motivates the following simple protocol for spin qudit tomography: select a random set of measurement axes $V$ by uniformly sampling points on the sphere \footnote{To sample a point $\p{\alpha,\beta}$ from the uniform distribution on the sphere (with azimuthal angle $\alpha$ and polar angle $\beta$), you can sample a point $\p{a,b}\in[0,1]\times[0,1]$ from the uniform distribution on the unit square, and then set $\alpha=2\pi a$ and $\beta=\arccos\p{1-2b}$.}, and optimize (using any standard minimization algorithm) the $2\abs{V}$ parameters in $V$ (two angles for each point $\v v\in V$) to minimize the quantum error scale $\epsilon_V^2$ defined in Eq.~\eqref{eq:scale}.
If $\abs{V}$ is too large for such optimization  to be practical, one can instead generate many random sets of measurement axes, and then choose the set with the smallest quantum error scale $\epsilon_V^2$.
After choosing a set of measurement axes $V$, (experimentally) make $n$ measurements of spin projection along every axis $\v v\in V$, and set $\tilde\Pi_{\v v\mu}\approx\bk{\Pi_{\v v\mu}}_\rho$ to the fraction of times in which the measurement outcome was $\mu$.
The estimates $\tilde\Pi_{\v v\mu}$ of $\bk{\Pi_{\v v\mu}}_\rho$ determine estimates of the expectation values $\bk{T_{\v v\ell,0}}_\rho$, which can be used to construct an estimate $\tilde\rho_V$ of $\rho$, defined in Eq.~\eqref{eq:reconstructed_state}.
If $\tilde\rho_V$ has negative eigenvalues, its distance from $\rho$ can be reduced with maximum-likelihood corrections \cite{smolin2012efficient}.

@@

  = \sum_{m,n} D^\ell_{m,0}\p{\v v_i}^* D^\ell_{n,0}\p{\v v_i}
  \cov_\rho\p{T_{\ell m}^\dag, T_{\ell n}},

@@

  = \f1n \sum_{i,\ell,m,n}

>>>>

Consider a $d$-state spin qudit with total spin $s\equiv\frac{d-1}{2}$.
The defining property of a spin qudit is the fact that it describes an angular momentum degree of freedom, which has specific implications for how a spin qudit should transform under the group SO(3) of rotations in 3D space.
Due to the central importance of these transformation rules for a spin qudit, we first identify a basis of operators that transforms nicely under 3D rotations.
One such basis is provided by the {\it transition operators} (also known as polarization operators \cite{kryszewski2006positivity, bertlmann2008bloch}), defined by
\begin{align}
  T_{\ell m} \equiv \sqrt{\f{2\ell+1}{2s+1}} \sum_{\mu,\nu=-s}^s
  \bk{s\mu;\ell m|s\nu} \op{\nu}{\mu},
  \label{eq:trans_op}
\end{align}
where $\bk{s\mu;\ell m|s\nu}$ is a Clebsh-Gordan coefficient that enforces $\ell\in\set{0,1,\cdots,d-1}$ and $m\in\set{-\ell,-\ell+1,\cdots,\ell}$, such that there are $d^2$ operators in total.
For brevity, we will generally treat the value of $d$ (or $s$) as constant but arbitrary throughout this work, and suppress any explicit dependence of quantities or operators such as $T_{\ell m}$ on $d$.
The transition operators are orthonormal with respect to the trace inner product, and transform nicely under conjugation:
\begin{align}
  \obk{T_{\ell m }|T_{\ell'm'}}
  = \delta_{\ell\ell'} \delta_{mm'},
  &&
  T_{\ell m}^\dag = \p{-1}^m T_{\ell,-m},
\end{align}
where for any $d\times d$ matrix $X=\sum_{\mu,\nu} X_{\mu\nu} \op{\mu}{\nu}$ we define the $d^2$-component vector $\oket{X} \equiv \sum_{\mu,\nu} X_{\mu\nu} \ket{\mu\nu}$; $\obra{X}$ is the conjugate transpose of $\oket{X}$, such that $\obk{X|Y}=\tr\p{X^\dag Y}$; and $\delta_{kk'}\equiv 1$ if $k=k'$ and $0$ otherwise.
These properties of the transition operators allow us to expand any density operator $\rho$ in the transition operator basis as
\begin{align}
  \rho = \sum_{\ell=0}^{d-1} \sum_{m=-\ell}^\ell
  \rho_{\ell m} T_{\ell m},
  &&
  \rho_{\ell m} \equiv \bk{T_{\ell m}^\dag}_\rho,
  \label{eq:trans_state}
\end{align}
where $\bk{X}_\rho\equiv\tr\p{\rho X}$, and $\rho^\dag=\rho$ implies that $\rho_{\ell m}^*=\p{-1}^m\rho_{\ell,-m}$.
The transition operators can be physically interpreted as the transition $\ket\psi\to T_{\ell m}\ket\psi$ induced on the state $\ket\psi$ of a spin-$s$ system by the absorption of a spin-$\ell$ particle with spin projection $m$ onto a fixed quantization axis.
In analogy with the spherical harmonics $Y_{\ell m}$, we will refer to the index $\ell$ as the {\it degree} and $m$ as the {\it order} of $T_{\ell m}$.

@@

Reconstructing an arbitrary density operator $\rho$ essentially requires us to find a set of axes $V=\set{\v v}$ and associated coefficients $C_{\ell mk}\p{\v v}$ that would allow us to recover any coefficient $\rho_{\ell m}$ in the expansion in Eq.~\eqref{eq:trans_state} through
\begin{align}
  \rho_{\ell m} = \bk{T_{\ell m}^\dag}_\rho
  = \sum_{\v v,k} C_{\ell mk}\p{\v v} \bk{T_{\v vk,0}^\dag}_\rho.
  \label{eq:state_recon}
\end{align}
Expanding the rotated transition operators $T_{\v vk,0}$ into a sum of un-rotated transition operators, according to Eq.~\eqref{eq:trans_rot}, we find that the recovery condition in Eq.~\eqref{eq:state_recon} is satisfied when
\begin{align}
  T_{\ell m}^\dag
  = \sum_{\v v,k,n} C_{\ell mk}\p{\v v}
  D^k_{n,0}\p{\v v}^* T_{kn}^\dag.
\end{align}
Orthogonality of the transition operators motivates the ansatz $C_{\ell mk}\p{\v v}=\delta_{k\ell}C_{\ell m}\p{\v v}$, which in turn implies that
\begin{align}
  \sum_{\v v} C_{\ell m}\p{\v v} D^\ell_{n,0}\p{\v v}^* = \delta_{mn}
  \label{eq:tomo_recovery}
\end{align}
for all $\ell$.

In fact, the problem of finding suitable axes $V$ and coefficients $C_{\ell m}\p{\v v}$ to satisfy Eq.~\eqref{eq:tomo_recovery}, solving which would enable the recovery of arbitrary spin qudit states, can be mapped onto the well-studied problem of signal recovery on the sphere \cite{mcewen2011novel, rauhut2011sparse, alem2012sparse, khalid2014optimaldimensionality}.
The signal recovery problem can be stated as follows: given a square-integrable function $f$ on the sphere, with the harmonic expansion
\begin{align}
  f\p{\v v} = \sum_{\ell,m} f_{\ell m} Y_{\ell m}\p{\v v},
\end{align}
where $f_{\ell m}$ are expansion coefficients and $Y_{\ell m}$ is the spherical harmonic of degree $\ell$ and order $m$, find a set of points $V=\set{\v v}$ and associated coefficients $\tilde C_{\ell m}\p{\v v}$ with which we can reconstruct $f$, or equivalently its coefficients $f_{\ell m}$, from knowledge of the function's value $f\p{\v v}$ at all points $\v v\in V$; that is
\begin{align}
  f_{\ell m} = \sum_{\v v} \tilde C_{\ell m}\p{\v v} f\p{\v v}
  = \sum_{\v v,k,n} \tilde C_{\ell m}\p{\v v} Y_{kn}\p{\v v} f_{kn}.
\end{align}
Reconstruction of functions with arbitrary coefficients $f_{\ell m}$ implies that
\begin{align}
  \sum_{\v v} \tilde C_{\ell m}\p{\v v} Y_{kn}\p{\v v}
  = \delta_{k\ell} \delta_{mn},
  \label{eq:full_recovery}
\end{align}
which is a stronger version of the condition that we found for the spin qudit tomography problem in Eq.~\eqref{eq:tomo_recovery}.
We will refer to Eq.~\eqref{eq:full_recovery} as the {\it full recovery problem}, and Eq.~\eqref{eq:tomo_recovery} as the the {\it reduced recovery problem}.
Due to the fact that $D^k_{n,0}\p{\v v}^* = \sqrt{\frac{4\pi}{2\ell+1}}\, Y_{kn}\p{\v v}$, any solution to the full recovery problem automatically solves the reduced recovery problem by setting $C_{\ell m}\p{\v v} = \sqrt{\frac{2\ell+1}{4\pi}}\, \tilde C_{\ell m}\p{\v v}$.
In principle, this mapping allows us to import a host of existing signal recovery algorithms \cite{mcewen2011novel, rauhut2011sparse, alem2012sparse, khalid2014optimaldimensionality} for the task of spin qudit tomography.
In practice, spin qudits typically have only a modest dimension $d$, which allows for simpler and optimized tomography protocols that are practical despite worse scaling with $d$ (see Section \ref{sec:protocol}).

@@

Since
\begin{enumerate*}
\item there are only $2\ell+1\le2d-1$ transition operators of degree $\ell\le\ell_{\t{max}}$, and
\item measuring spin projection along any single axis provides measurement data for transition operators of {\it all} degrees $\ell\le\ell_{\t{max}}$,
\end{enumerate*}
one might expect that $2d-1$ suitably chosen measurement axes would always suffice to perform full tomography of a $d$-level spin qudit.
In the language of classical signal recovery, we would expect there to exist a fixed set of $2\ell_{\t{max}}+1$ points on the sphere that enable the recovery of {\it any} function with fixed degree $\ell\le\ell_{\t{max}}$ (given prior knowledge of $\ell$) \footnote{We could further conjecture that if {\it any} set of points $V\subset\Omega_2$ enables the recovery of degree-$\ell_{\t{max}}$ functions on the sphere, then $V$ enables the recovery degree-$\ell$ functions with $\ell\le\ell_{\t{max}}$.  However, this condition is stronger than necessary to guarantee full recovery of a $d$-level spin qudit state with $2d-1$ measurement axes, which only requires the existence of {\it some} such set of points.}.
Unfortunately, we have no general proof of this conjecture, which requires a deeper analysis of fixed-degree signal recovery on the sphere.
Nonetheless, in the following section we numerically verify that $2d-1$ measurement axes suffice for all $d\le200$ [\red{todo: possibly refer to Section \ref{sec:axes} instead}].

@@

Computing the classical error scale $\S_V^2$ to certify the bound in Eq.~\eqref{eq:bound_eps} requires constructing the measurement matrix $M_V$ and finding its singular values.
The complexity of this task can be greatly reduced by the fact that the degree $\ell$ of a transition operator $T_{\ell m}$ is preserved under rotations, which implies that the unitary
\begin{align}
  U \equiv \sum_{\ell=0}^{d-1} \sum_{m=-\ell}^\ell
  \oket{T_{\ell m}} \bra{\ell m},
  \label{eq:trans_basis}
\end{align}
with vectors $\oket{T_{\ell m}}$ in a column indexed by integers $\p{\ell,m}$, block-diagonalizes the measurement matrix into $d$ blocks indexed by the degree $\ell$:
\begin{align}
  U^\dag M_V = \sum_{\ell=0}^{d-1} \op{\ell} \otimes M_{V\ell},
\end{align}
where the $\p{2\ell+1}\times\abs{V}$ blocks are
\begin{align}
  M_{V\ell} \equiv \sum_{m,\v v} \ket{m} \obk{T_{\ell m}|T_{\v v\ell,0}} \bra{\v v}
  = \sum_{m,\v v} D^\ell_{m,0}\p{\v v} \op{m}{\v v}.
\end{align}
As the singular values of $M_V$ are invariant under unitary transformations, it follows that
\begin{align}
  \S_V^2 = \sum_\ell \S_{V\ell}^2,
  &&
  \S_{V\ell}^2 \equiv \norm{M_{V\ell}^{-1}}^2,
\end{align}
where $M_{V\ell}^{-1}$ is the right inverse of $M_{V\ell}$.
Constructing the block $M_{V\ell}$ and computing its singular values takes at most $O(\abs{V}d^2)$ time.
If we assume that $\abs{V}\sim d$, then computing $\S_V^2$ to certify the bound in Eq.~\eqref{eq:bound_eps} takes $O(d^4)$ serial or $O(d^3)$ parallel runtime (see Figure \ref{fig:times}).

\begin{figure}
  \centering
  \includegraphics{error_scale/error_times.pdf}
  \caption{Serial runtime to compute the error scales $\S_V^2$ and $\epsilon_V^2$ with $\abs{V}=2d-1$ randomly chosen measurement axes.
    Each point is an average over 100 calculations or 5 minutes of runtime, whichever comes first.
    Dashed lines shows a fit to runtime $t=c d^\alpha$ for all $d\ge100$, finding $\alpha\approx 3.7$.
    These results do not include fixed $O(d^3)$ runtimes for pre-computing quantities that can be recycled for every new choice of measurement axes $V$.}
  \label{fig:times}
\end{figure}

The assumption that observables can be estimated up to uncorrelated noise with maximal variance $\epsilon^2$, summarized by Eq.~\eqref{eq:error_assumption}, is reasonable when measurement error is dominated by classical sources of experimental noise.
However, this assumption breaks down when measurement error is limited by fundamental quantum (shot) noise.
We relax the assumption of Eq.~\eqref{eq:error_assumption} in Appendix \ref{sec:error}, where we instead assume that $\tilde\rho_V$ is built from $n$ independent measurements of spin projection along every axis $\v v\in V$, with shot noise the dominant source of error.
In this case, the mean squared distance between the approximation $\tilde\rho_V$ and the true state $\rho$ can be compactly expressed using trace inner products of the {\it covariance matrices}
\begin{align}
  \C_\ell\sp{\rho} \equiv \sum_{m,n=-\ell}^\ell
  \cov_\rho\p{T_{\ell m}^\dag, T_{\ell n}} \op{m}{n}
\end{align}
where $\cov_\rho\p{X,Y} \equiv \bk{XY}_\rho - \bk{X}_\rho \bk{Y}_\rho$, and the {\it noise matrices}
\begin{align}
  \N_{V\ell} \equiv M_{V\ell} \diag\sp{M_{V\ell}^{-1}
    \p{M_{V\ell}^{-1}}^\dag} M_{V\ell}^\dag,
\end{align}
where $\diag\sp{X}$ sets all off-diagonal parts of $X$ to zero:
\begin{align}
  \Bbk{\norm{\tilde\rho_V-\rho}^2}
  = \f1n \sum_\ell \obk{\N_{V\ell}|\C_\ell\sp{\rho}}.
  \label{eq:error}
\end{align}
Maximizing this error over all $\rho$ (for a fixed set of axes $V$) yields a reconstruction error bound of the form
\begin{align}
  \Bbk{\norm{\tilde\rho_V-\rho}^2} \le \f{\epsilon_V^2}{n},
  \label{eq:bound}
\end{align}
where the complexity of computing the {\it quantum error scale} $\epsilon_V^2$ (defined below) is the same as that of computing the classical error scale $\S_V^2$ (see Figure \ref{fig:times}).
For comparison with the bound in Eq.~\eqref{eq:bound_eps}, we note that $\epsilon_V^2<\S_V^2/2$, so the previous bound still holds with the replacement $\epsilon^2\to1/2n$.
We also note that the (squared) reconstruction error in Eq.~\eqref{eq:error} and its bound in Eq.~\eqref{eq:bound} obey the ``standard quantum limit'' of $\sim1/\sqrt{n}$ scaling in the number of measurements.
In principle, this scaling can be improved to $\sim1/n$ by preparing and measuring entangled copies of many qudits \cite{giovannetti2006quantum}.

@@

The remainder of this section is devoted to defining $\epsilon_V^2$, which is derived in Appendix \ref{sec:error}, and can be skipped without consequence for the discussions in the rest of this paper.
To write out the quantum error scale $\epsilon_V^2$ explicitly, we define a matrix of structure constants
\begin{align}
  g_{L\ell} \equiv \sum_{m,n=-\ell}^\ell
  \obk{T_{L,m+n}|T_{\ell m}T_{\ell n}} \op{m}{n},
\end{align}
and the inversion operator $\I\equiv\sum_m\p{-1}^m\op{-m}{m}$, which are used to define the $\p{2\ell+1}$-component vectors
\begin{align}
  \ket{\chi_{VL}}
  &\equiv \sum_{M=-L}^L \chi^V_{LM} \ket{M}, \\
  \chi^V_{Lm} &\equiv \sum_\ell \obk{\N_{V\ell}|\D_M|\I g_{L\ell}}, \\
  \D_M &\equiv \sum_{m,m'} \delta_{M,m'-m} \op{mm'}.
\end{align}
Here $\D_M$ simply picks off the $M$-th diagonal of the matrix it acts on, such that $\obk{\N_{V\ell}|\D_M|\I g_{L\ell}}$ is an inner product between the $M$-th diagonals of $\N_{V\ell}$ and $\I g_{L\ell}$.
We then define
\begin{align}
  \L_{V\ell} \equiv \N_{V\ell} + \I\N_{V\ell}\I,
\end{align}
in terms of which we have that $\epsilon_V^2 \equiv \sum_\ell \epsilon_{V\ell}^2$, with
\begin{align}
  \epsilon_{V\ell}^2 \equiv
  \bk{\chi_{V\ell}|\p{\L_{V\ell}^{-1}
      - \L_{V\ell}^{-1}\N_{V\ell}\L_{V\ell}^{-1}}|\chi_{V\ell}}.
\end{align}

@@

The ability to certify a statistical error bound on the empirical estimate $\tilde\rho_V$ of an unknown quantum state $\rho$ motivates the following simple protocol for spin qudit tomography: select a random set of measurement axes $V$ by uniformly sampling points on the sphere \footnote{To sample a point $\p{\alpha,\beta}$ from the uniform distribution on the sphere (with azimuthal angle $\alpha$ and polar angle $\beta$), you can sample a point $\p{a,b}\in[0,1]\times[0,1]$ from the uniform distribution on the unit square, and then set $\alpha=2\pi a$ and $\beta=\arccos\p{1-2b}$.}, and optimize (using any standard minimization algorithm) the $2\abs{V}$ parameters in $V$ (two angles for each point $\v v\in V$) to minimize the quantum error scale $\epsilon_V^2$.
If $\abs{V}$ is too large for such optimization  to be practical, one can instead generate many random sets of measurement axes, and then choose the set with the smallest quantum error scale $\epsilon_V^2$.
After choosing a set of measurement axes $V$, (experimentally) make $n$ measurements of spin projection along every axis $\v v\in V$, and set $\tilde\Pi_{\v v\mu}\approx\bk{\Pi_{\v v\mu}}_\rho$ to the fraction of times in which the measurement outcome was $\mu$.
The estimates $\tilde\Pi_{\v v\mu}$ of $\bk{\Pi_{\v v\mu}}_\rho$ determine estimates of the expectation values $\bk{T_{\v v\ell,0}}_\rho$, which can be used to construct an estimate $\tilde\rho_V$ of $\rho$.
If $\tilde\rho_V$ has negative eigenvalues, its distance from $\rho$ can be reduced with maximum-likelihood corrections \cite{smolin2012efficient}.

@@

  = \sum_{m,m'} D^\ell_{m,0}\p{\v v_i}^* D^\ell_{m',0}\p{\v v_i}
  \cov_\rho\p{T_{\ell m}^\dag, T_{\ell m'}},

@@

  = \f1n \sum_{i,\ell,m,m'}
